Towards atomistic resolution structure of phosphatidylcholine headgroup and glycerol backbone at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}

### 2015-10-11 HGmodelMANUSCRIPT/REPLYtoREVIEWER2.txt

We thank for the very positive feedback. The errors in English are
now fixed as suggested.

Reviewer 1.

1.
Author reply:

2.
Author reply:

All the "subjective sweet spots" in Fig. 2 have exactly the same size: 0.04 units.

This size was agreed on based on the high experimental repeatibility of the order
parameters that we discuss in the text (p12): "Almost all the reported values are
within a variation of ±0.02 [...] for all fully hydrated PC bilayers, regardless
of variation in their acyl chain composition and temperature."

To make the (equal) size of the sweet spots clear to the reader, we now explicitly
point out this size when introcuding them in the text (p15):

"Motivated by the high experimental reproducibility, we have highlighted in Fig. 2
subjective sweet spots (light blue areas **spanning 0.04 units**), within which
we expect the calculated values of the order parameters of a well-performing
force field to fall."

The added words are those between the stars (**).

### 2015-10-11 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Tuned language.

The order parameters of the glycerol backbone and headgroup calculated from different force fields for various lipids have been 
previously compared to experiments~\cite{shinoda97,hogberg08,castro08,klauda10,kapla12,dickson12,poger12,ferreira13,chowdhary13,maciejewski14}. 
The general conclusion from these studies seems to be that the CHARMM based~\cite{hogberg08,klauda10}, GAFFlipid~\cite{dickson12} and
MacRog~\cite{maciejewski14} force fields perform better for the glycerol backbone and headgroup structures than the GROMOS based models~\cite{castro08,kapla12,poger12,ferreira13}.
However, none of the studies exploits the full potential of the available experimental data discussed in previous section, i.e. the quantitative accuracy, known signs and stereospecific labeling of
the experimental order parameters.

To get a general idea of the quality of the glycerol backbone and choline headgroup structures in different models, we calculated 
the order parameters for these parts from thirteen different lipid models (Table~\ref{systems}) and 
plotted the results together with experimental values in Fig.~\ref{HGorderparameters}.
Two criteria were used to judge the quality of the model: there must not be significant  {\bf forking} in the $\alpha$ and $\beta$ carbons,
there must be only moderate forking in the g$_3$ carbon and there must be significant forking in the g$_1$ carbon, the {\bf magnitude}
should be preferably inside to the subjective sweet spots determined from experiments (blue shaded regions in Fig.~\ref{HGorderparameters}).
The results for each force field with respect to the above criteria are summarized in Figure~\ref{FullHydrationComparisonTable}.
\begin{figure}[]
  \centering
  \includegraphics[width=8.6cm]{../DATAreportediINblog/comparisonTable.pdf}
  \newline
  \caption{\label{FullHydrationComparisonTable}
Rough subjective ranking of force fields based on Fig.~\ref{HGorderparameters}.
%
Here "M" indicates a magnitude problem, "F" a forking problem;
letter size increases with problem severity.
%
Color scheme:
"within experimental error" (dark green),
"almost within experimental error" (light green),
"clear deviation from experiments" (light red), and
"major deviation from experiments" (dark red).
%
The $\Sigma$-column shows the total deviation of the force field,
when individual carbons are given weights of 0 (matches experiment), 1, 2, and 4 (major deviation).
%
For full details of the assessment, see Supporting Information.
} 
\end{figure}

None of the studied force fields fulfils these criteria completely, however CHARMM36 is close. 
This is not surprising since the dihedral potentials in this model are tuned to reproduce these 
parameters better against experiments~\cite{klauda10}.
The next models in the list are CHARMM36-UA~\cite{henin08,lee14} and H\"ogberg08~\cite{hogberg08} which is also not surprising since
these models are using CHARMM bonded potentials for glycerol backbone and choline. The fourth and the fifth models in the list, MacRog~\cite{maciejewski14} and
GAFFlipid~\cite{dickson12}, have independently determined dihedral potentials. All the models based on Gromos potentials and Slipids perform less well.
In the present work we subject the CHARMM36, MacRog, GAFFlipid and Berger-POPC-07 to a more careful comparison including the sterospecific labeling  
(Fig.~\ref{HGorderparameters2}), and atomistic level structure and responses to the dehydration and cholesterol content in the following sections.
These models are selected for more detailed studies since they are the best representatives of different dihedral potential parametrization techniques 
(CHARMM36, MacRog, GAFFlipid), and the Berger based models are the most used lipid models in the literature.

>>>>

The order parameters of the glycerol backbone and headgroup calculated from different force fields for various lipids have been 
previously compared to experiments~\cite{shinoda97,hogberg08,castro08,klauda10,kapla12,dickson12,poger12,ferreira13,chowdhary13,maciejewski14}. 
The general conclusion from these studies seems to be that the CHARMM based~\cite{hogberg08,klauda10}, GAFFlipid~\cite{dickson12} and
MacRog~\cite{maciejewski14} force fields perform better for the glycerol backbone and headgroup structures than the GROMOS based models~\cite{castro08,kapla12,poger12,ferreira13}.
However, none of the studies exploits the full potential of the available experimental data discussed in the previous section, i.e. the quantitative accuracy, known signs and stereospecific labeling of
the experimental order parameters.

To get a general idea of the quality of the glycerol backbone and choline headgroup structures in different models, we calculated 
the order parameters for these parts from thirteen different lipid models (Table~\ref{systems}) and 
plotted the results together with experimental values in Fig.~\ref{HGorderparameters}.
Two criteria were used to judge the quality of the model: 1) there must not be significant  {\bf forking} in the $\alpha$ and $\beta$ carbons,
there must be only moderate forking in the g$_3$ carbon and there must be significant forking in the g$_1$ carbon, and 2) the {\bf magnitude}
should be preferably inside to the subjective sweet spots determined from experiments (blue shaded regions in Fig.~\ref{HGorderparameters}).
The results for each force field with respect to the above criteria are summarized in Figure~\ref{FullHydrationComparisonTable}.
\begin{figure}[]
  \centering
  \includegraphics[width=8.6cm]{../DATAreportediINblog/comparisonTable.pdf}
  \newline
  \caption{\label{FullHydrationComparisonTable}
Rough subjective ranking of force fields based on Fig.~\ref{HGorderparameters}.
%
Here "M" indicates a magnitude problem, "F" a forking problem;
letter size increases with problem severity.
%
Color scheme:
"within experimental error" (dark green),
"almost within experimental error" (light green),
"clear deviation from experiments" (light red), and
"major deviation from experiments" (dark red).
%
The $\Sigma$-column shows the total deviation of the force field,
when individual carbons are given weights of 0 (matches experiment), 1, 2, and 4 (major deviation).
%
For full details of the assessment, see Supporting Information.
} 
\end{figure}

None of the studied force fields fulfils these criteria completely, however CHARMM36 is close. 
This is not surprising since the dihedral potentials in this model are tuned against experiments to better reproduce these 
order parameters~\cite{klauda10}.
The next models in the list are CHARMM36-UA~\cite{henin08,lee14} and H\"ogberg08~\cite{hogberg08}, which is also not surprising since
these models are using the CHARMM bonded potentials for glycerol backbone and choline. The fourth and the fifth models in the list, MacRog~\cite{maciejewski14} and
GAFFlipid~\cite{dickson12}, have independently determined dihedral potentials. All the models based on Gromos potentials and Slipids perform less well.
In the following sections we subject CHARMM36, MacRog, GAFFlipid and Berger-POPC-07 to a more careful comparison including the stereospecific labeling  
(Fig.~\ref{HGorderparameters2}), atomistic level structure, and responses to dehydration and cholesterol content.
These models are selected for more detailed study since they are the best representatives of different dihedral potential parametrization techniques 
(CHARMM36, MacRog, GAFFlipid), and the Berger based models are the most used lipid models in the literature.
